\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-11-06 demystify.tex
updates

\subsection{Research or Personal Statement}\label{sec:research-statement}

@@

While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
statement! So write it well because it could make a big difference.
In your statement, you have the opportunity to make  make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
%is important because if you need (GTA) funding, it will provide evidence
%that you can teach and communicate with students.

@@

\begin{commentbox}[Hung:]
I think asking for a “research vision” from a Ph.D. applicant is too much. Even people graduated with a Ph.D. have a great difficulty in explicating their researh vision. In my opinion, you should focus on showing (not telling) your research passion (if you do have).
\end{commentbox}

@@

TA is common in the beginning when you haven't found an adviser who would pay you RA. As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation). Your TA is paid through the school/department, i.e., they hire you to help teach.  During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding is not typically available during the summer, which has few courses.

@@

Choosing a school and an adviser is clearly one of the most important things in your mind when you apply to and especially when you get admitted.  This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.  This section aims to mitigate some confusions and help you make informed decision.

@@

We will first discuss about universities in the US that offer PhD in CS.

@@

    In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants? 
    \tcblower
    \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities. Many probably will say no as they do not have the bandwidth to take on random strangers, but some may say yes if they see you have a matching profile.  
\end{commentbox}

@@

@@

\subsubsection{PhD in other related fields: CE, IST, Cybersecurity}

@@

>>>>

\subsection{Statement of Purpose (SOP)}\label{sec:research-statement}

@@

While you might not have control over LORs or \hyperref[sec:your-school]{where your go to school}, you do over your
statement of purpose (SOP) or personal statement! So write it well because it could make a big difference.
In your statement, you have the opportunity to make  make your application stand out and unique, which can make you \emph{fit} the CS program you're applying to, even if you don't have very strong research experience.
A well-written LOR also shows that you can communicate, which is very important in research, and that you can effectively teach and communicate with students, which is important for TA funding (see \S\ref{sec:funding}).
%is important because if you need (GTA) funding, it will provide evidence
%that you can teach and communicate with students.

@@

\begin{commentbox}[Hung:]
I think asking for a “research vision” from a Ph.D. applicant is too much. Even people graduated with a Ph.D. have a great difficulty in explicating their research vision. In my opinion, you should focus on showing (not telling) your research passion (if you do have).
\end{commentbox}

@@

TA is common in the beginning when you haven't found an adviser who would pay you RA. As a TA, you spend up to 20 hrs/week and help with classes (e.g., grading or teaching labs/recitation). Your TA is paid through the school or department, i.e., they hire you to help teach.  During a semester, a TA might work with several courses and professors (not necessary their adviser).  TA funding is not typically available during the summer, which has few courses.

@@

Choosing a school and an adviser is clearly one of the most important things in your mind when you apply and especially when you get admitted.  This is further complicated due to cultural differences and unfamiliarity of  international students to the US higher education system.  This section aims to mitigate some confusions and help you make informed decision.

@@

We will first discuss about universities in the US that offer PhD in CS. Then we will talk about how to rank and select them.

@@

    In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants? 
    \tcblower
    \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities. Many probably will say no as they do not have the bandwidth to take on random students, but some may say yes if they see potential fits.  
\end{commentbox}

@@

It is worth noting that professors in a non-CS department also unlikely can serve as formal PhD advisers for CS students (see \S\ref{sec:related-fields}).

@@

\subsubsection{PhD in other related fields: CE, IST, Cybersecurity}\label{sec:related-fields}

@@

\paragraph{About Me} I am an assistant professor at George Mason University (GMU). Prior to GMU, I was at the University of Nebraska-Lincoln (UNL). I have been involved in the PhD admission process at GMU and UNL for many years and currently I am the program director of the Masters program in Software Engineering at GMU (thus also have some experience with the MS admission process--which is quite different than PhD). I got my undergraduate and MS degrees from Penn State. I did my PhD at the University of New Mexico-Albuquerque, and then a postdoc at the University of Maryland-College Park.  I had also worked at the Naval Research Lab and Lockheed Martin.

While I grew up in the US, I am familiar with international students, e.g., many of my students are international students.  I have many international collaborators and also have been a mentor for many students from Vietnam.  I also have many colleague and friends who were international students and have heard their stories.  I hope to capture their challenges and experiences in this document so that it can be useful for prospective students.
